---
title: "Models with Machine Learning"
author: "vaibhavi"
date: "March 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plots of crime trends and machine learning models  
```{r}
setwd("~/Projects/Detroit Project/detroit-crime-education/data/clean")
crime = read.csv("zip-level-encoded-crime.csv")
```

## Plotting Trends in Crime across the Years (Time)
You can also embed plots, for example:
```{r}
crime

#Aggregate over Years to find Clustering Model
yearbackup = aggregate(crime,by=list(crime$YEAR), FUN= sum)
yearbackup

drops= c("X", "YEAR","ZIP_CHAR")
yearbackup = yearbackup[, !names(yearbackup) %in% drops]
yearbackup

colnames(yearbackup)
library(data.table)
setnames(yearbackup, old = c('Group.1'), new = c('YEAR'))
yearbackup

```

## Plotting Trends in Crime across ZIPS (Spatial)
```{r}
crime

#Aggregate over Years to find Clustering Model
zipbackup = aggregate(crime,by=list(crime$ZIP_CHAR), FUN= sum)
zipbackup

drops= c("X", "YEAR","ZIP_CHAR")
zipbackup = zipbackup[, !names(zipbackup) %in% drops]
zipbackup

setnames(zipbackup, old = c('Group.1'), new = c('ZIP'))
zipbackup

```
## To do later!!!
```{r}
library(ggplot2)
library(reshape2)
d <- melt(yearbackup, id.vars="YEAR")

# Everything on the same plot
ggplot(d, aes(year,value, col=variable)) + 
  geom_point() + 
  stat_smooth() 

# Separate plots
ggplot(d, aes(year, value)) + 
  geom_point() + 
  stat_smooth() +
  facet_wrap(~variable)

```
##Analyze Zip Level Data
Most types of crimes seem to be correlated with high corr values!!!
The neighborhoods that have crime, have all kinds of crime!
```{r}

zipbackup
colnames(zipbackup)

keeps = ""
zipshort = zipbackup[,c("ZIP","NUM_SUBSTANCE","NUM_WHITECOLLAR", "NUM_PROPERTY_RELATED","NUM_VIOLENT") ]
plot(zipshort)

cor(zipshort$NUM_SUBSTANCE, zipshort$NUM_WHITECOLLAR)

The
```

##Cluster Analysis
http://www.sthda.com/english/wiki/cluster-analysis-in-r-unsupervised-machine-learning

```{r}
# Install factoextra and cluster
#install.packages("factoextra")
#install.packages("cluster")

library(factoextra)
library(cluster)
```

## Let the cool visualization of distance between zips begin!
```{r}

scalezip = zipshort

scalezip$ZIP = zipshort$ZIP
rownames(scalezip) = scalezip$ZIP
rownames(zipshort) = zipshort$ZIP

scalezip$NUM_SUBSTANCE = scale(zipshort$NUM_SUBSTANCE) #Don't scale ZIP
scalezip$NUM_WHITECOLLAR = scale(zipshort$NUM_WHITECOLLAR)
scalezip$NUM_VIOLENT = scale(zipshort$NUM_VIOLENT)
scalezip$NUM_PROPERTY_RELATED = scale(zipshort$NUM_PROPERTY_RELATED)

scalezip
colnames(scalezip)
rownames(scalezip) = scalezip[,"ZIP"]

scalezip

res.dist = get_dist(scalezip, stand =TRUE, method="pearson")
fviz_dist(res.dist, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

#res.dist
res.dist2 = get_dist(scalezip, stand =TRUE, method="euclidean")
fviz_dist(res.dist2, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

```
#Let's do K-means clustering
How many clusters do we need? *8*
```{r}
fviz_nbclust(zipshort, kmeans, method = "gap_stat")

```


```{r}
km.res <- kmeans(zipshort, 5, nstart = 25)

# Visualize
fviz_cluster(km.res, data = zipshort, frame.type = "convex")+ theme_minimal()
```




```{r}
pam.res <- pam(zipshort, 5)
# Visualize
fviz_cluster(pam.res)
```

```{r}
set.seed(20)
zipCluster <- kmeans(zipshort[, 2:5], 5, nstart = 20)
zipCluster

```
```{r}


zips <- zipshort$ZIP
zipCluster$cluster

set1 = zips[zipCluster$cluster==1]
set2 = zips[zipCluster$cluster==2]
set3 = zips[zipCluster$cluster==3]
set4 = zips[zipCluster$cluster==4]
set5 = zips[zipCluster$cluster==5]

set1

zipshort$CLUSTER = 1
zipshort$CLUSTER = ifelse(zipshort$ZIP %in% set1, 1,  
                          ifelse(zipshort$ZIP %in% set2, 2,  
                                 ifelse(zipshort$ZIP %in% set3, 3,  
                                        ifelse(zipshort$ZIP %in% set4, 4, 5) ) ) )
zipshort

zipshort$CLUSTER = ifelse(zipshort$CLUSTER ==1, "EXTREME",  
                          ifelse(zipshort$CLUSTER ==2, "HIGH",  
                                 ifelse(zipshort$CLUSTER ==3, "SAFE",  
                                        ifelse(zipshort$CLUSTER ==4, "MEDIUM", "LOW") ) ) )

zipshort$CLUSTER = as.factor(zipshort$CLUSTER)
levels(zipshort$CLUSTER)
zipshort$CLUSTER = factor(zipshort$CLUSTER, levels = c("SAFE", "LOW", "MEDIUM","HIGH","EXTREME"))

levels(zipshort$CLUSTER)
zipshort

```

##Save Files for Further Analysis
```{r}
write.csv(zipshort, "zip-level-crime-clusters.csv")
write.csv(yearbackup, "year-level-crimes.csv")
```



##Using GGPLOT ot Visualize Cluster
```{r}
head(zipshort)
#Already cluster is a factor, no need to convert
#zipshort$cLUSTER <- as.factor(zipshort$cLUSTER)
ggplot(data=zipshort, aes(NUM_WHITECOLLAR, NUM_VIOLENT, color = CLUSTER)) + geom_point()

ggplot(data=zipshort, aes(NUM_PROPERTY_RELATED, NUM_SUBSTANCE, color = CLUSTER)) + geom_point()


```

Visualizing R!
https://cran.r-project.org/web/packages/broom/vignettes/kmeans.html

```{r}
set.seed(2014)
points <-zipshort %>% group_by(CLUSTER) %>%
    do(data.frame(NUM_WHITECOLLAR=rnorm(.$size[1], .$NUM_WHITECOLLAR[1]),
                  x2=rnorm(.$size[1], .$x2[1])))

library(ggplot2)
ggplot(points, aes(x1, x2, color=cluster)) + geom_point()
```

#Doing PCA & K-Means (NOT DONE)
```{r}
zipshort
# log transform 
log.zipshort <- log(zipshort[, 2:5])
zipcodes <- zipshort$ZIP
clusters = zipshort$CLUSTER 

# apply PCA - scale. = TRUE is highly 
# advisable, but default is FALSE. 


zip.pca <- prcomp(log.zipshort,
                 center = TRUE,
                 scale. = TRUE) 


```



```{r}



```
