---
title: "Crime-Category-Encoded"
author: "vaibhavi"
date: "March 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the data

```{r}
setwd("~/Projects/Detroit Project")
crime=read.csv("crimewithzips.csv")
```

## Keeping relevant data

```{r}
colnames(crime)
drops=c("X","CASEID", "ADDRESS", "SCA", "CENSUSTRACT")

crime = crime[, !names(crime) %in% drops]
colnames(crime)

#crime[crime$CATEGORY=="ESCAPE",] #Fugitive
#crime[crime$CATEGORY=="ENVIRONMENT",] #Runaway (Teenagers?)
#crime[crime$CATEGORY=="KIDNAPING",]
#crime[crime$CATEGORY=="OBSTRUCTING JUDICIARY",] # Parole Violation
crime[crime$CATEGORY=="OBSTRUCTING THE POLICE",] #Resisting Arrest
```


## Adding category-specific crime  columns
Top 15 crime categories are:
ASSAULT
LARCENY
BURGLARY
DAMAGE TO PROPERTY
STOLEN VEHICLE

AGGRAVATED ASSAULT
MURDER/INFORMATION
TRAFFIC
FRAUD
ROBBERY

DANGEROUS DRUGS
WEAPONS OFFENSES
ESCAPE
OBSTRUCTING JUDICIARY
DISORDERLY CONDUCT

NUM_CRIMES column must be added as well

```{r}
crime$NUM_CRIMES = 1

# Top 15 crimes in Wayne County
crime$NUM_ASSAULT <- ifelse(crime$CATEGORY == "ASSAULT", 1, 0)
crime$NUM_LARCENY <- ifelse(crime$CATEGORY == "LARCENY", 1, 0)
crime$NUM_BURGLARY <- ifelse(crime$CATEGORY == "BURGLARY", 1, 0)
crime$NUM_PROPERTY <- ifelse(crime$CATEGORY == "DAMAGE TO PROPERTY", 1, 0)
crime$NUM_AUTOTHEFT <- ifelse(crime$CATEGORY == "STOLEN VEHICLE", 1, 0)

crime$NUM_AGG_ASSAULT <- ifelse(crime$CATEGORY == "AGGRAVATED ASSAULT", 1, 0)
crime$NUM_MURDER <- ifelse(crime$CATEGORY =="MURDER/INFORMATION", 1, 0)
crime$NUM_TRAFFIC <- ifelse(crime$CATEGORY == "TRAFFIC", 1, 0)
crime$NUM_FRAUD <- ifelse(crime$CATEGORY == "FRAUD", 1, 0)
crime$NUM_ROBBERY <- ifelse(crime$CATEGORY == "ROBBERY", 1, 0)

crime$NUM_DRUGS <- ifelse(crime$CATEGORY == "DANGEROUS DRUGS", 1, 0)
crime$NUM_WEAPONS <- ifelse(crime$CATEGORY == "WEAPONS OFFENSES", 1, 0)
crime$NUM_ESCAPE <- ifelse(crime$CATEGORY == "ESCAPE", 1, 0)
crime$NUM_OBSTRUCT <- ifelse(crime$CATEGORY == "OBSTRUCTING JUDICIARY", 1, 0)
crime$NUM_CONDUCT <- ifelse(crime$CATEGORY == "DISORDERLY CONDUCT", 1, 0)

# Violent Crimes
crime$NUM_VIOLENT <- ifelse(crime$CATEGORY %in% c("HOMICIDE","NEGLIGENT HOMICIDE", "JUSTIFIABLE HOMICIDE", "MURDER/INFORMATION", "SEX OFFENSES", "AGGRAVATED ASSAULT", "ASSAULT", "KIDNAPING"), 1, 0)


# Property
crime$NUM_PROPERTY <- ifelse(crime$CATEGORY %in% c("ARSON","LARCENY", "DAMAGE TO PROPERTY", "STOLEN PROPERTY", "OTHER BURGLARY", "ROBBERY", "STOLEN VEHICLE", "BURGLARY"), 1, 0)


# Youth-Related Crimes
crime$NUM_YOUTH <- ifelse(crime$CATEGORY %in% c("LARCENY","DRUNKENNESS", "LIQUOR", "OUIL", "DAMAGE TO PROPERTY", "STOLEN PROPERTY", "OBSCENITY", "DISORDERLY CONDUCT","DANGEROUS DRUGS", "ASSAULT", "VAGRANCY (OTHER)","RUNAWAY","OBSTRUCTING THE POLICE", "TRAFFIC OFFENSES", "TRAFFIC", "ENVIRONMENT"), 1, 0)

# White Collar Crime
crime$NUM_WHITECOLLAR <- ifelse(crime$CATEGORY %in% c("FORGERY","FRAUD", "BRIBERY", "EMBEZZLEMENT", "EXTORTION"), 1, 0)

# Substance Abuse Related Crime
crime$NUM_SUBSTANCE <- ifelse(crime$CATEGORY %in% c("DRUNKENNESS", "LIQUOR", "OUIL","DANGEROUS DRUGS"), 1, 0)


colnames(crime)

```
## Save the Incident Level Data
1211903 rows, 35 columns
```{r}
head(crime)
colnames(crime)
dim(crime)
write.csv(crime, "encodedcrimeincidents.csv")
colnames(crime)
```



## Sanity Check Before Aggregating to Per Year, Per Zip data

579 rows, 22 columns
```{r}
head(crime)
colnames(crime)
#Drop any columns that cannot be grouped for Yearly  Zip data
drops = c("ROWNUM", "CATEGORY", "OFFENSEDESCRIPTION", "INCIDENTDATE","HOUR","PRECINCT","COUNCIL","NEIGHBORHOOD","LON","LAT", "LOCATION", "MONTH", "FINALZIP")
backup = crime[ , ! names(crime) %in% drops]

#22 columns
colnames(backup)
head(backup)
backup = aggregate(backup, by=list(backup$YEAR, backup$ZIP_CHAR), FUN=sum)

head(backup)
backup$YEAR <- NULL
backup$ZIP_CHAR <-NULL
#library(data.table)
setnames(backup, old = c('Group.1','Group.2'), new = c('YEAR','ZIP_CHAR'))
backup

dim(backup)
```


## Plotting Data (To be Done Later)
```{r}
#library(ggplot2)
#ggplot(backup, aes(YEAR,value, col=variable)) + 
 # geom_point() + 
  #stat_smooth() 
```



## Saving the  Aggregate data
```{r}
write.csv(backup, "zip-level-encoded-crime.csv")
```
