---
title: "merge-crime-education.Rmd"
author: "vaibhavi"
date: "March 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Examining Crime with Zips Data frame 
Year, ZIP, No of crimes, No of schools in district, Mean Graduation Rate, Mean Dropput rate, Total graduates, total drop outs
```{r}
#setwd("~/Projects/Detroit Project")
crime <- read.csv("crimewithzips.csv")
colnames(crime)
```
[1] "X"                  "ROWNUM"             "CASEID"             "ADDRESS"           
[5] "CATEGORY"           "OFFENSEDESCRIPTION" "INCIDENTDATE"       "HOUR"              
[9] "SCA"                "PRECINCT"           "COUNCIL"            "NEIGHBORHOOD"      
[13] "CENSUSTRACT"        "LON"                "LAT"                "LOCATION"          
[17] "MONTH"              "YEAR"               "FINALZIP"           "ZIP_CHAR"   


Crime Colnames to discard for final merged model:
------------------------------------------
X, Rownum, CaseId, Address, Precinct, OffenseDescription, IncidentDate, Hour, Category*, Neighbourhood, SCA
Censustract, Lat, Lon, Location, Month, FinalZip(has errors)

Crime Colnames to keep for merged model:
----------------------------------
Year, ZipChar, Number of crimes


Crime Colnames to add for merged model:
---------------------------------
Number of violent crimes, number of non-violent, number of  white-collar/ blue-collar crimes
IsViolent? Is Non-Violent? Is it white-collar? Is it Blue-collar? etc


## Creating simplified Crime model
```{r}
keep = c("YEAR", "ZIP_CHAR")
simple.crime=crime[keep]
colnames(simple.crime)

simple.crime$NUM_CRIMES = 1
colnames(simple.crime)
head(simple.crime)

simple.crime=aggregate(simple.crime$NUM_CRIMES, by=list(simple.crime$YEAR, simple.crime$ZIP_CHAR), FUN=sum)
colnames(simple.crime)

colnames(simple.crime) = c("YEAR", "ZIP", "NUM_CRIMES")

head(simple.crime)
summary(simple.crime)
simple.crime


plot(simple.crime)

#Let's see how crimes vary across the years
plot(simple.crime$NUM_CRIMES ~simple.crime$YEAR)

#Wouldn't it be better as a violin plot?
library(ggplot2)

#Change the input X variable (YEAR) to a factor from int
simple.crime$YEAR = as.factor(simple.crime$YEAR)

# Change violin plot line colors by groups
p<-ggplot(simple.crime, aes(x=simple.crime$YEAR, y=simple.crime$NUM_CRIMES, color=simple.crime$YEAR)) +
  geom_violin(trim=FALSE)
p + stat_summary(fun.y=median, geom="point", size=2, color="red") + stat_summary(fun.y=mean, geom="point", shape=23, size=2)


# Change box plot line colors by groups
p<-ggplot(simple.crime, aes(x=simple.crime$YEAR, y=simple.crime$NUM_CRIMES, color=simple.crime$YEAR)) +  geom_boxplot()
p


#Re-order number of crimes per year , with zips



```


## Examining Education with Zips Data frame 
Year, ZIP, No of crimes, No of schools in district, Mean Graduation Rate, Mean Dropput rate, Total graduates, total drop outs, 
```{r}
#setwd("~/Projects/Detroit Project/detroit-crime-education/data/clean/Education")
education <- read.csv("clean-education-with-zips.csv")
colnames(education)
```
 
 The returned Column Names are:
 [1] "X"                "ENROLL_YEAR"      "SCHOOL_NAME"      "COHORT_GRAD_YEAR" "GRAD_YEAR"        "NUM_COHORT"      
 [7] "NUM_GRADUATES"    "NUM_DROPOUTS"     "GRAD_RATE"        "DROP_RATE"        "ZIP"              "LATITUDE"        
[13] "LONGITUDE"
 

Education Colnames to discard for merged model:
------------------------------------------


Education Colnames to keep for merged model:
----------------------------------
GRAD_YEAR, COHORT_GRAD_YEAR, NUM_DROPOUTS, NUM_GRADUTES


Education Colnames to add for merged model:
---------------------------------
Number of Schools 


## Creating simplified Education model
```{r}

colnames(education)
education$NUM_SCHOOLS= 1
colnames(education)

#Can remove individual school identifiers like school name, latitude and longitude
#Need to look at Zip code level data per year
education  #615 rows

#Having aggregated data for Enroll Year and Zip Code
keep1= c("ENROLL_YEAR", "COHORT_GRAD_YEAR","NUM_SCHOOLS","NUM_COHORT", "NUM_GRADUATES", "NUM_DROPOUTS", "ZIP")
edu.enroll = education[keep1]
edu.enroll=aggregate(edu.enroll, by=list(edu.enroll$ENROLL_YEAR, edu.enroll$ZIP), FUN=sum)
edu.enroll
head(edu.enroll)  #421 rows

#Something weird in the last two rows :( Check out why the zip code is coming as 96480, must be some doubling of the data!
colnames(edu.enroll) 
filter=c("Group.1", "Group.2","NUM_SCHOOLS","NUM_COHORT","NUM_GRADUATES", "NUM_DROPOUTS")
edu.enroll=edu.enroll[filter]
colnames(edu.enroll)= c("ENROLL_YEAR", "ZIP","NUM_SCHOOLS","NUM_COHORT","NUM_GRADUATES", "NUM_DROPOUTS")
edu.enroll$COHORT_GRAD_YEAR = edu.enroll$ENROLL_YEAR +4

head(edu.enroll)
edu.enroll

#Add columns for dropout and graduation rates
edu.enroll$GRAD_RATE = edu.enroll$NUM_GRADUATES/edu.enroll$NUM_COHORT
edu.enroll$DROP_RATE = edu.enroll$NUM_DROPOUTS/edu.enroll$NUM_COHORT

head(edu.enroll)
edu.enroll


#Having aggregated data for Graduation Year and Zip Code
keep2= c("GRAD_YEAR","NUM_SCHOOLS","NUM_COHORT", "NUM_GRADUATES", "NUM_DROPOUTS", "ZIP")
edu.grad= education[keep2]
edu.grad=aggregate(edu.grad, by=list(edu.grad$GRAD_YEAR, edu.grad$ZIP), FUN=sum)
edu.grad #332 rows

#Last Record is weird!!!!! Must check.
colnames(edu.grad) 
filter=c("Group.1", "Group.2","NUM_SCHOOLS","NUM_COHORT","NUM_GRADUATES", "NUM_DROPOUTS")
edu.grad=edu.grad[filter]
colnames(edu.grad)= c("GRAD_YEAR", "ZIP","NUM_SCHOOLS","NUM_COHORT","NUM_GRADUATES", "NUM_DROPOUTS")

head(edu.grad)
edu.grad

#Add columns for dropout and graduation rates
edu.grad$GRAD_RATE = edu.grad$NUM_GRADUATES/edu.grad$NUM_COHORT
edu.grad$DROP_RATE = edu.grad$NUM_DROPOUTS/edu.grad$NUM_COHORT

head(edu.grad)
edu.grad

```

## Write down the simplified zip level education data into a csv
```{r}
write.csv(edu.enroll,"zip_level_education_enroll.csv")
write.csv(edu.grad,"zip_level_education_grad.csv")


simple.crime
write.csv(simple.crime,"zip_level_crime.csv")


```



## Merging the two simple models for education and crime!
```{r}
simple.crime$YEAR <- as.numeric(as.character(simple.crime$YEAR))
edu_crime <- merge(edu.grad, simple.crime,by.x=c("GRAD_YEAR","ZIP"),by.y=c("YEAR","ZIP"),all.x = TRUE)
simple.crime <- data.table(simple.crime)
edu_crime$NUM_CRIMES[is.na(edu_crime$NUM_CRIMES)] <- 0
edu_crime

write.csv(edu_crime,"edu-crime-merged.csv")
```


## Machine Learning! Regression Models
```{r}
#TBD
```
