---
title: "merge-crime-education.Rmd"
author: "vaibhavi"
date: "March 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Examining Crime with Zips Data frame 
Year, ZIP, No of crimes, No of schools in district, Mean Graduation Rate, Mean Dropput rate, Total graduates, total drop outs
```{r}
#setwd("~/Projects/Detroit Project")
crime <- read.csv("crimewithzips.csv")
colnames(crime)
```
[1] "X"                  "ROWNUM"             "CASEID"             "ADDRESS"           
[5] "CATEGORY"           "OFFENSEDESCRIPTION" "INCIDENTDATE"       "HOUR"              
[9] "SCA"                "PRECINCT"           "COUNCIL"            "NEIGHBORHOOD"      
[13] "CENSUSTRACT"        "LON"                "LAT"                "LOCATION"          
[17] "MONTH"              "YEAR"               "FINALZIP"           "ZIP_CHAR"   


Crime Colnames to discard for final merged model:
------------------------------------------
X, Rownum, CaseId, Address, Precinct, OffenseDescription, IncidentDate, Hour, Category*, Neighbourhood, SCA
Censustract, Lat, Lon, Location, Month, FinalZip(has errors)

Crime Colnames to keep for merged model:
----------------------------------
Year, ZipChar, Number of crimes


Crime Colnames to add for merged model:
---------------------------------
Number of violent crimes, number of non-violent, number of  white-collar/ blue-collar crimes
IsViolent? Is Non-Violent? Is it white-collar? Is it Blue-collar? etc


## Creating simplified Crime model
```{r pressure, echo=FALSE}
keep = c("YEAR", "ZIP_CHAR")
simple.crime=crime[keep]
colnames(simple.crime)

simple.crime$NUM_CRIMES = 1
colnames(simple.crime)
head(simple.crime)

simple.crime=aggregate(simple.crime$NUM_CRIMES, by=list(simple.crime$YEAR, simple.crime$ZIP_CHAR), FUN=sum)
colnames(simple.crime)

colnames(simple.crime) = c("YEAR", "ZIP", "NUM_CRIMES")

head(simple.crime)
summary(simple.crime)

plot(simple.crime)
```


## Examining Education with Zips Data frame 
Year, ZIP, No of crimes, No of schools in district, Mean Graduation Rate, Mean Dropput rate, Total graduates, total drop outs, 
```{r}
setwd("~/Projects/Detroit Project/detroit-crime-education/data/clean/Education")
education <- read.csv("Education_Data_wLoc.csv")
colnames(education)
```

 [1] "Location.Name"          "Rate.Year"              "Cohort.Graduation.Year" "Total.Cohort"          
 [5] "Total.Graduated"        "OffTrack.Other"         "Dropouts"               "Graduation.Rate"       
 [9] "Dropout.Rate"           "gradyear"               "enrollyear"             "School.Name"           
 [13] "ZIP.Code"               "Latitude"               "Longitude" 
 

Education Colnames to discard for merged model:
------------------------------------------
Location.Name, Total.Cohort, Rate.Year

Education Colnames to keep for merged model:
----------------------------------
Cohort.Graduation.Year, Total.Cohort, Total.Graduated, OffTrack.Other, Dropouts, Graduation.Rate, Dropout.Rate, 


Education Colnames to add for merged model:
---------------------------------
Number of Schools 


## Creating simplified Education model
```{r}
#keep = c("YEAR", "ZIP_CHAR")
#simple.crime=crime[keep]
#colnames(simple.crime)

#simple.crime$NUM_CRIMES = 1
#colnames(simple.crime)
#head(simple.crime)

#simple.crime=aggregate(simple.crime$NUM_CRIMES, by=list(simple.crime$YEAR, simple.crime$ZIP_CHAR), FUN=sum)
#colnames(simple.crime)

#colnames(simple.crime) = c("YEAR", "ZIP", "NUM_CRIMES")

#head(simple.crime)
#summary(simple.crime)

#plot(simple.crime)
```

Problem: 
Inflated datasets!!!!

## Merging the two simple models for education and crime!
```{r}

```


## Machine Learning! Regression Models
```{r}

```
