---
title: "Time Series"
author: "Shrawan Sher"
date: "3/5/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ggplot2)
library(lubridate)
library(reshape) 
library(data.table)
library(astsa)
library(forecast)

#Fetching the crime data 
crime <- read.csv("crimewithzips.csv")
colnames(crime)

#Selecting just month and year columns
crimebytime <- crime[,c(17:18)]

#Adding a dummy column for crime count
crimebytime$DUMMY <- 1
head(crimebytime)
```


```{r}
#Casting the data across Year and month 
tseries <- dcast(crimebytime, YEAR~MONTH,fun.aggregate = sum,value.var = "DUMMY")

#Converting the cast into a vector
tsr <-as.vector(t(tseries[,2:13]))

#Removing crime data for 2017
tsr <- tsr[1:96]

#Time series till Jun 2016
ts2017 <- tsr[1:90]
ts2017 <- ts(ts2017,frequency = 12, start=c(2009,1))

#Creating a time series till Dec 2016
times <- ts(tsr,frequency=12,start=c(2009,1))

plot(times,main="Number of crimes reported 2009-2016" ,ylab="Number of Crimes",xlab="Month-Year")
```




#Decomposition of the time series
comp <- decompose(times)
decomp <- stl(times,s.window = "periodic")
plot(comp)
plot(decomp)

#Remove the seasonality from the time series
deseasonalcrime <- seasadj(decomp)
plot(deseasonalcrime)


#Stationarity
# formal statistical test for stationarity
library(tseries)
adf.test(times, alternative = "stationary")
Acf(times,main='')
Pacf(times,main='')
#Spikes at 1, 13 for pacf 
auto.arima(times, seasonal=FALSE)


#Can we trust this model? We can start by examining ACF and PACF plots for model residuals. 
#If model order parameters and structure are correctly specified, 
#we would expect no significant autocorrelations present. 

fit<-auto.arima(times, seasonal=FALSE)
tsdisplay(residuals(fit), lag.max=45, main='(1,1,1) Model Residuals')

#There is a clear pattern present in ACF/PACF and model residuals plots repeating at lag 12. 
#This suggests that our model may be better off with a different specification, 
#such as p = 12 or q = 12. 

fit2 = arima(times, order=c(12,1,1))

fit2

tsdisplay(residuals(fit2), lag.max=15, main='Seasonal Model Residuals')


#Forecasting

fcast <- forecast(fit2, h=6)
plot(fcast)
library(zoo)
#pd<-funggcast(y,fcast)
ggplot()

#When are crimes most likely
crimecat <- crime[,c(6,17:18)]



#Where do crimes occur


#Machine Learning
