---
title: "education-cutdown"
author: "vaibhavi"
date: "March 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Examining Education with Zips Data frame 
Year, ZIP, No of crimes, No of schools in district, Mean Graduation Rate, Mean Dropput rate, Total graduates, total drop outs, 
```{r}
setwd("~/Projects/Detroit Project/detroit-crime-education/data/clean/Education")
education <- read.csv("Education_Data_wLoc.csv")
colnames(education)


head(education)
```

Remove Columns: Rate.Year, Graduation Rate, Dropout Rate, 

#Create dataframe:
SchoolName, enrollyear, gradyear, ZipCode, Latitude, Longtitude
TotalCohort, Total Graduated, Dropouts

How to reduce from 4 year, 5 year, 6 year program to 1 record?
Total.Cohort (Max)
Total.Graduated (Max)
Dropouts(Max)

Then Add columns:
grad rate(calculate)
dropout rate(calculate)
offtrack(calculate)

```{r}

#Dropping is not meaningful for factors, hence converting factors to chars
head(education)
education$School.Name = as.character(education$School.Name)
education$Location.Name = as.character(education$Location.Name)
education$Rate.Year = as.character(education$Rate.Year)
write.csv("Education_Data_wLoc.csv")

colnames(education)

drops <- names(education) %in% c("Rate.Year", "OffTrack.Other","Graduation.Rate", "Dropout.Rate", "Location.Name") 
small.education <- education[!drops]
#dim(small.education) /#Same 1426 rows
  
#Convert factors to characters
head(small.education)

clean.education=aggregate(small.education, by=list(small.education$enrollyear, small.education$ZIP.Code, small.education$School.Name), FUN=max)
head(clean.education)

#Re-order columns
#clean.education = clean.education["enrollyear"," "," "]

#rename columns
colnames(clean.education)

#Save to file

```


```{r}

```


```{r}

```

