---
title: "education-cutdown"
author: "vaibhavi"
date: "March 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The aim is to cut down the Rate-Year column,and have only one entry per year for a school.


## Examining Education with Zips Data frame 
Year, ZIP, No of crimes, No of schools in district, Mean Graduation Rate, Mean Dropput rate, Total graduates, total drop outs, 
```{r}
#setwd("~/Projects/Detroit Project/detroit-crime-education/data/clean/Education")
education <- read.csv("Education_Data_wLoc.csv")
colnames(education)


head(education)
```

Remove Columns: Rate.Year, Graduation Rate, Dropout Rate, 

#Create dataframe:
SchoolName, enrollyear, gradyear, ZipCode, Latitude, Longtitude
TotalCohort, Total Graduated, Dropouts

How to reduce from 4 year, 5 year, 6 year program to 1 record?
Total.Cohort (Max)
Total.Graduated (Max)
Dropouts(Max)

Then Add columns:
grad rate(calculate)
dropout rate(calculate)
offtrack(calculate)

```{r}

#Dropping is not meaningful for factors, hence converting factors to chars
head(education)
education$School.Name = as.character(education$School.Name)
education$Location.Name = as.character(education$Location.Name)
education$Rate.Year = as.character(education$Rate.Year)
write.csv("Education_Data_wLoc.csv")

colnames(education)

drops <- names(education) %in% c("Rate.Year", "OffTrack.Other","Graduation.Rate", "Dropout.Rate", "Location.Name") 
small.education <- education[!drops]
#dim(small.education) /#Same 1426 rows
  
#Convert factors to characters
head(small.education)

clean.education=aggregate(small.education, by=list(small.education$enrollyear, small.education$ZIP.Code, small.education$School.Name), FUN=max)
head(clean.education)

#Re-order columns, and remove"Group.1,2,3" columns added by aggregate
keep =c("enrollyear", "School.Name", "Cohort.Graduation.Year", "gradyear", "Total.Cohort", "Total.Graduated", "Dropouts", "ZIP.Code", "Latitude", "Longitude") 
clean.education= clean.education[keep]

#rename columns
colnames(clean.education) =c("ENROLL_YEAR","SCHOOL_NAME", "COHORT_GRAD_YEAR", "GRAD_YEAR","NUM_COHORT","NUM_GRADUATES", "NUM_DROPOUTS","ZIP","LATITUDE","LONGITUDE")

#Add Graduation Rate & Dropout Rate
clean.education$GRAD_RATE = clean.education$NUM_GRADUATES/clean.education$NUM.COHORT
clean.education$DROP_RATE = clean.education$NUM_DROPOUTS/clean.education$NUM.COHORT

#Re-order columns, and move newly added columns before ZIP & Lat, Long
keep =c( "ENROLL_YEAR", "SCHOOL_NAME", "COHORT_GRAD_YEAR", "GRAD_YEAR","NUM_COHORT","NUM_GRADUATES","NUM_DROPOUTS","GRAD_RATE"        ,"DROP_RATE","ZIP" ,"LATITUDE","LONGITUDE") 
clean.education= clean.education[keep]
colnames(clean.education)

```


We have now reduced to 615 rows only, and have recalculated the graduation and drop out rates per school per enrollment year.
All variables are int or double variables. No Factors!

```{r}

#Verify one last time
dim(clean.education)
head(clean.education)


#Save to file
write.csv(clean.education, "clean-education-with-zips.csv")

```

